Reconstructing Propositional Proofs in Type Theory

### 2017-10-24 paper/sections/introduction.tex
[ paper ] added the fourth paragraph in the introduction.

Reconstructing and verifying proofs delivered by automatic theorem provers
(henceforth ATP) are non-trivial tasks.
In the last decades, \ATPs is fast becoming a key instrument in 
different disciplines (\eg verifying cryptographic protocols for information security) and researchers have shown an increased interest to prove the validity of the deduction algorithms.
One reason to say that are the bugs found in many ATPs.
For that reason, in order to give confidence to the user of these systems,
many ATPs has started to include in their outputs, the proofs of the
conclusion. However, existing research recognizes that in many cases these proofs encode non-trivial reasoning
hard to reconstruct them and therefore to verify them.

@@

One approach to address the proof-reconstruction 
is integrating the \ATP, the \emph{source} system, with a proof assistant, the \emph{target} system.
The target system becomes in the proof checker.   
Previous studies have reported proof-reconstruction using proof assistant for classical logic where the development is at a mature stage.
In contrast, we choose a proof assistant for type theory
where proof verification becomes a type-checking task of the
proof-term following the Curry-Howard isomorphism.

% To guarantee the validity of their methods,
% many ATPs has been opted to include proofs in their output.

% The problem address in this paper concern about justifying such proofs.

% One approach to address  these tasks could be use a proof assistant
% to verify the reply of the ATP.

% % During the last decades, the research field of automated reasoning has been experimenting a growth in the development of new methods, tools, and systems. One of the major achievements of this field are the automatic
% % theorem provers (henceforth ATPs).

% At the beginning in its development, the main goal was prove automatically
% problems in different logics (\eg, first-order logic, high-order logic).
% Nowadays, they continuing extending its support, performance and precision
% in their deductive methods.

% Nevertheless, it has been argued that some of their deductions are not
% valid even in delivering a proof.

% This problem to verify proofs delivered by ATPs is a non-trivial task
% and involves an in-depth understanding of the algorithms in the \ATP.
% The proof-reconstruction is a non-trivial problem.

% The proof-reconstruction is a non-trivial task since it depends on the
% integration between two complex systems.

@@

% One way to reconstruct proofs of \ATPs consists of using an \ITP.
% An \ITP system provide us the logic framework to check and validate
% the reply of the \ATPs since they allow us to define the formal language
%  for the problems \ie, operators, logic variables, axioms, and theorems.

@@

% Since the formalism of the source (the evidence
% generated by the \ATP) is not necessarily the same logic in the
% target, the reconstruction turns out to be a ``reverse engineering''
% task.

@@

% The recontruction of  proof can play an important role in addressing a issue
% of ...
% Moreover, verification is fast becoming a key instrument in automatic reasoning.

% The type theory provides a useful account of how ...

% No previous study has investigated to reconstruct proofs using a syntactical approach
% in \Agda for any \ATPs.

In this paper, we ported a subset of the \Metis inference rules to
\Agda, the propositional fragmented. The formalization of
these rules allowed us to justify step-by-step \Metis derivations by
translating automatically to \Agda by a the \Athena tool.

>>>>

Reconstructing and verifying proofs delivered by automatic theorem provers
(henceforth ATP) are non-trivial tasks.
In the last decades, \ATPs is fast becoming a key instrument in 
different disciplines (\eg verifying cryptographic protocols for information security) and researchers have shown an increased interest to prove the validity of the deduction algorithms.
One reason to say that is the bugs found in many ATPs~\cite{Keller2013,Bohme2011,Fleury2014}.
For that reason, in order to give confidence to the user of these systems,
many ATPs has started to include in their outputs, the proofs of the
conclusion. However, existing research recognizes that in many cases these proofs encode non-trivial reasoning
hard to reconstruct them and therefore to verify them.

@@

One approach to address the proof-reconstruction 
is integrating the \ATP, the \emph{source} system, with a proof assistant, the 
\emph{target} system.
The target system becomes in the proof checker. These proof assistants allow 
us to define the formal language used in the proofs, \ie, operators, logic 
variables, axioms, and theorems. A proof-reconstruction tool provides such an
integration, translating the replied generated by the prover into the formalism of the proof assistant.

Previous studies have reported proof-reconstruction using proof assistants for classical logic where the development is at a mature stage.
In contrast, we choose a proof assistant for type theory
where proof verification becomes a type-checking task of the proof-term following the Curry-Howard isomorphism. Then, the proof-reconstruction generates such proof-terms by reconstructing the proof replied by the \ATP.
However, some limitations arises by using type theory as we describe in Section~\ref{sec:limitations-type-theory}.

In this paper, we describe how to reconstruct proofs in type theory
replied by \Metis Prover\footnote{v2.3 release 2017102}.
We ported the subset of the \Metis inference rules for the propositional 
fragment using a syntactical treatment of that logic. The \Metis reasoning was formalized in \Agda~\cite{AgdaProp,AgdaMetis}. 
to reconstruct \Metis proofs, we built a proof-reconstruction tool named
\Athena~\cite{Athena} that is able to generate \Agda\footnote{v2.5.3} proof-terms for \Metis derivations. We have avoided the use of propositions meanings towards a future work to support other formalisms.
% but also to contrast the deep embedding using to write propositions in \Agda 
% with a shallow embedding.

@@

% One way to reconstruct proofs of \ATPs consists of using an \ITP.
% An \ITP system provide us the logic framework to check and validate
% the reply of the \ATPs since

@@

@@

In this paper,

The formalization of these rules allow us to justify step-by-step \Metis derivations. We implemented this formalization in \Agda and we built
a translator tool, \Athena, to generate automatically the proof-terms
from \Metis proofs.
